Setting parameters for biological models with ANIMO

### 2014-04-02 Paper.tex
Added files for presentation

% anche Related work: es. Copasi
% Altro su parameter sweep/sensitivity analysis? Noi usiamo un approccio un po' naif, senza tante pretese, senza tanti numeri.
%parametric model checking, hytech?
%imitator II ??
\section{Conclusions}
ANIMO has been extended with a proper support for parameter synthesis. Now the tool can be used more efficiently
to model biological networks, without concentrating exclusively on their topology.
In particular, the addition of an automatic parameter sweep feature allows the user
to save considerable amounts of time, keeping closer to experimental data without the need
to perform a hand-made parameter search.

In the future, we plan to improve our parameter analysis techniques, possibly using some of
the techniques already developed for testing parameter sensitivity~\cite{inverse-ta1, inverse-ta2} and
robustness~\cite{robust-ta1, robust-ta2, robust-ta3, robust-ta4} in \tas\ and hybrid systems more in general~\cite{decidability-ta-ha, imitator2}.
Techniques adopted in other formalisms such as e.g. differential equations~\cite{copasi, breach, kinfer}, rule-based languages~\cite{biocham, rulebender}
or Bayesian networks~\cite{bayes-infer} are also being
considered as a source of inspiration.
As already stated, parameter sensitivity is very important in biological models: if a treatment
derived from a model is based on narrow bounds for a parameter, that parameter must not exhibit significant
variety among patients. By introducing an automatic analysis for parameter sensitivity, our tool
could point out the zones of a network model where the user should concentrate more, for example by
improving the topology, or by more focused parameter sweeps.
Moreover, we also aim at applying some results from automata learning~\cite{test-based-modelling} to the generation of
\tas\ models based on experimental data: generating models that exhibit specific behaviors
implies a careful definition of their parameter space.
Finally, parametric model checking~\cite{param-mc} represents an interesting development direction:
having a tool automatically output the parameter intervals for which some
given outputs are observed is a highly valued resource for the biological researcher.
% Investigating the properties of the specific \tas\ models used in ANIMO
% could lead to important tractability results valid in our case, allowing us
% to apply methods not valid for \tas\ in general.
%FUTURE WORK!!
%parameter sensitivity: se un trattamento dipende da un parametro specifico, e' il caso che non si registri troppa varieta' per quel parametro tra i pazienti.
% per quei parametri non sensibili, si eviti di concentrare troppi sforzi. Il tool potrebbe automaticamente determinare le ``zone calde'' della rete
%anche alla luce di una possibile generazione di modelli t.a. basati su dati sperimentali, trovare modi intelligenti di generare i parametri
%il fatto che i modelli che usiamo possano identificare un sottinsieme ristretto di t.a. potrebbe essere vantaggioso per la trattabilita' dei problemi
%parametrized model checking: con che parametri e' possibile raggiungere questo stato?
%definisci un obiettivo di quello che vuoi ottenere, e lui ti dice i [range di] parametri che arrivano a quell'evoluzione
% Possibile future work: vedere meglio altri modi di applicare parameter setting nel contesto dei t.a., usare
% calcolo distribuito per accelerare la ricerca dei parametri

>>>>

% anche Related work: es. Copasi
% Altro su parameter sweep/sensitivity analysis? Noi usiamo un approccio un po' naif, senza tante pretese, senza tanti numeri.
%parametric model checking, hytech?
%imitator II ??
\section{Conclusions}
ANIMO has been extended with a proper support for parameter synthesis. Now the tool can be used more efficiently
to model biological networks, without concentrating exclusively on their topology.
In particular, the addition of an automatic parameter sweep feature allows the user
to save considerable amounts of time, keeping closer to experimental data without the need
to perform a parameter search by hand.

In the future, we plan to improve our parameter analysis techniques, possibly using some of
the techniques already developed for testing parameter sensitivity~\cite{inverse-ta1, inverse-ta2} and
robustness~\cite{robust-ta1, robust-ta2, robust-ta3, robust-ta4} in \tas\ and hybrid systems more in general~\cite{decidability-ta-ha, imitator2}.
Techniques adopted in other formalisms such as e.g. differential equations~\cite{copasi, breach, kinfer}, rule-based languages~\cite{biocham, rulebender}
or Bayesian networks~\cite{bayes-infer} are also being
considered as a source of inspiration.
As already stated, parameter sensitivity is very important in biological models: if a treatment
derived from a model is based on narrow bounds for a parameter, that parameter must not exhibit significant
variety among patients. By introducing an automatic analysis for parameter sensitivity, our tool
could point out the zones of a network model where the user should concentrate more, for example by
improving the topology, or by performing more focused parameter sweeps.
Moreover, we also aim at applying some results from automata learning~\cite{test-based-modelling} to the generation of
\tas\ models based on experimental data: generating models that exhibit specific behaviors
implies a careful definition of their parameter space.
Finally, parametric model checking~\cite{param-mc} represents an interesting development direction:
having a tool automatically output the parameter intervals for which some
given outputs are observed is a highly valued resource for the biological researcher.
% Investigating the properties of the specific \tas\ models used in ANIMO
% could lead to important tractability results valid in our case, allowing us
% to apply methods not valid for \tas\ in general.
%FUTURE WORK!!
%parameter sensitivity: se un trattamento dipende da un parametro specifico, e' il caso che non si registri troppa varieta' per quel parametro tra i pazienti.
% per quei parametri non sensibili, si eviti di concentrare troppi sforzi. Il tool potrebbe automaticamente determinare le ``zone calde'' della rete
%anche alla luce di una possibile generazione di modelli t.a. basati su dati sperimentali, trovare modi intelligenti di generare i parametri
%il fatto che i modelli che usiamo possano identificare un sottinsieme ristretto di t.a. potrebbe essere vantaggioso per la trattabilita' dei problemi
%parametrized model checking: con che parametri e' possibile raggiungere questo stato?
%definisci un obiettivo di quello che vuoi ottenere, e lui ti dice i [range di] parametri che arrivano a quell'evoluzione
% Possibile future work: vedere meglio altri modi di applicare parameter setting nel contesto dei t.a., usare
% calcolo distribuito per accelerare la ricerca dei parametri
